---
title: "Chi-Square Goodness-of-Fit"
author: "Patrice Lewis and Julissa Martinez"
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    toc: true
    code-fold: true
---

# What is Chi-Square Goodness-of-Fit

The Chi-square goodness-of-fit test is used to assess whether an observed distribution differs from the expected distribution among categorical variables in discrete data. This test measures the extent to which a statistical model fits the observations.

## Assumptions

-   The average expected counts are \> or = '5'
-   Categorical variables
-   All individual expected counts are \> or = '1' and independent of each other
-   The sample was randomly selected

# R Function

chisq.test(x, p)

x: a numeric vector for observed frequencies p: a numeric vector for expected proportions

NB: This test exist in the 'stats' package

# Load Libraries

```{r setup}
#| warning: false
#| code-fold: false

library(tidyverse)
library(table1)
library(skimr)
library(ggplot2)
```

# Example 1: Performing Chi-Square Goodness-of-Fit Test

Data source: "titanic.csv"

```{r data-download}
#| warning: false
#| code-fold: false

titanic <- read_csv("raw_data/titanic.csv")

```

## Hypothesis 1

H0: Passenger classes were not equally common on-board the Titanic\
H1: Passenger classes were equally common on-board the Titanic

## Data Exploration & Cleaning

```{r data-cleaning}
#| warning: false
#| code-fold: false

# Uses skim to get descriptive statistics of the data
skim(titanic)

# str(titanic)

# Examines variable of interest - passenger class
summary(titanic$Pclass)

# Three distinct classes - 1,2,3
unique(titanic$Pclass)

# Selects variables of interest
titanic_cleaned <- titanic %>% 
  select(
    pid = PassengerId,
    pclass = Pclass,
  )

# Outputs frequency table of passenger class
pclass_table <-
  table1( 
    ~ factor(pclass), 
    data = titanic_cleaned
)

pclass_table
```

## Graphs

```{r graphs}
#| warning: false


# Outputs pclass histogram to visualize frequency and distribution
hist_pclass <- titanic_cleaned %>% 
pull(pclass) %>% 
  hist(main = "Passenger class onboard Titanic")

# Data slightly right-skewed


#### Bar chart ####

ggplot(
  titanic_cleaned,
  aes(x = pclass,
      fill = pclass
  )
) +
  labs(
    x = "Passenger class",
    y = "Frequency"
  ) +
  scale_fill_viridis_d(option = "magma") +
  geom_bar()



```

## Chi-square Test

```{r pclass-chisq-test}
#| warning: false
#| code-fold: false

p_class <- c(216, 184, 491)
props <- c(0.242, 0.207, 0.551)

test_pclass <- chisq.test(p_class, p = props )

test_pclass

# Printing the p-value
test_pclass$p.value


# Printing the statistic
test_pclass$statistic



```

At a 0.05 level of significance, there is no sufficient evidence to show that passenger classes were equally common on-board the Titanic.
